---
title: "Titanic 2"
output: html_notebook
---

# Load the data and libraries
```{r, message = FALSE}
library('caret')
library('dplyr')
library('e1071')
train <- read.csv('./input/train.csv', stringsAsFactors = F, na.strings="")
test  <- read.csv('./input/test.csv', stringsAsFactors = F, na.strings="")
```
# Look at training Samples
```{r}
summary(train)
```
```{r}
str(train)
```
```{r}
colSums(is.na(train))
```
## Notes:
  1. Survival rate in training set is 0.3838384 which is greater than the actual survival rate of 31.6.  May need to subsample for class imbalance.
  2. There are 177 NA's in Age, as well as most of cabin, 2 embarked
  3. With only 891 obs will likely use some sort of linear regression

# Create new features
### Title
```{r}
createNewFeatures <- function(data) {
  trainMod <- data
  trainMod$Title <- gsub('(.*, )|(\\..*)', '', data$Name)
  
  rare_title <- c('Dona', 'Lady', 'the Countess','Capt', 'Col', 'Don', 'Dr', 'Major', 'Rev', 'Sir', 'Jonkheer')
  trainMod$Title[trainMod$Title == 'Mlle']        <- 'Miss'
  trainMod$Title[trainMod$Title == 'Ms']          <- 'Miss'
  trainMod$Title[trainMod$Title == 'Mme']         <- 'Mrs' 
  trainMod$Title[trainMod$Title %in% rare_title]  <- 'Rare Title'
  trainMod$Embarked[is.na(trainMod$Embarked)] <- "S"
  return(trainMod)
}
trainMod <- createNewFeatures(train)
testMod <- createNewFeatures(test)
```
# Train
```{r}
factorize <- function(data) {
  colsToFactor <- c("Pclass", "Sex", "Parch", "SibSp", "Title", "Embarked")
  data[colsToFactor] <- lapply(data[colsToFactor], factor)
  return(data)
}

preProcImp <- function(data) {
  preProcValues <- preProcess(data, method="knnImpute")
  trainModImp <- predict(preProcValues, data)
  return(trainModImp)
}

trainMod <- factorize(trainMod)
trainMod$Survived <- factor(trainMod$Survived)
testMod <- factorize(testMod)

trainModImp <- preProcImp(trainMod)
testModImp <- preProcImp(testMod)
testModImp
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10)

fitControlGlmBoost <- trainControl(method = "repeatedcv",
                           number = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary,
                           repeats = 10)

glmModel <- train(Survived ~ Pclass + Sex + Age + SibSp + Fare + Title + Embarked, data = trainModImp, 
                 method = "glm",
                 trControl = fitControl)
summary(glmModel)
```

```{r}
#gbmModel <- train(Survived ~ Pclass + Sex + Age + SibSp + Fare + Title, data = trainModImp, 
 #                method = "gbm",
  #               trControl = fitControl)

drops = c("Name", "Ticket", "Cabin", "Embarked", "PassengerId")
glmBoostPreProc <- function(data) {
  levels(data$Title) <- make.names(levels(data$Title))
  return(data[, -which(names(data) %in% drops)])
}

glmBoostTest <- glmBoostPreProc(testModImp)
glmBoostTrain <- glmBoostPreProc(trainModImp)
str(glmBoostTrain)

glmBoostModel <- train(Survived ~ ., data = glmBoostTrain[,-which(names(glmBoostTrain) %in% drops)], 
                 method = "glmboost", metric="ROC", tuneLength=5,
                 # family=binomial,
                 family=Binomial(link = c("logit")),
                 trControl = fitControlGlmBoost)

rf_model<-train(Survived~., data = trainModImp[,-which(names(trainModImp) %in% drops)],method="rf",
                trControl=trainControl(method="cv",number=5),
                prox=TRUE,allowParallel=TRUE)

#make.names(levels(trainModImp$Survived))
#str(trainModImp[,-which(names(trainModImp) %in% drops)])

summary(gbmModel)
summary(glmModel)
summary(glmBoostModel)
plot(glmBoostModel)
plot(rf_model)
plot(gbmModel)
```
# Predict glm
```{r}
prediction <- predict(glmModel, testModImp, type = "raw")

solution <- data.frame(PassengerID = test$PassengerId, Survived = prediction)

solution
write.csv(solution, file = 'glm_solution_3.csv', row.names = F)
```

# Predict gbm
```{r}
predictionGbm <- predict(gbmModel, testModImp, type = "raw")

solutionGbm <- data.frame(PassengerID = test$PassengerId, Survived = prediction)

solution

write.csv(solution, file = 'gbm_solution_1.csv', row.names = F)
```

# Predict glmBoost
```{r}
glmBoostTest = glmBoostTest[ifelse(glmBoostTest$Parch == "7", "6", glmBoostTest$Parch) , ]
str(glmBoostTest)
predictionGlmBoostm <- predict(glmBoostModel, glmBoostTest)

solutionGlmBoost <- data.frame(PassengerID = test$PassengerId, Survived = prediction)

solutionGlmBoost

write.csv(solutionGlmBoost, file = 'glm_boost_1.csv', row.names = F)
```


# Predict rf
```{r}
#predictionGlmBoostm <- predict(rf_model, testModImp[,-which(names(testModImp) %in% drops)])
testModImpRF = testModImp[ifelse(testModImp$Parch == "7", "6", testModImp$Parch) , ]
predictionRF <- predict(rf_model, testModImpRF)
solutionRF <- data.frame(PassengerID = test$PassengerId, Survived = prediction)

solutionRF

write.csv(solutionRF, file = 'rf_2.csv', row.names = F)
```


# Predict with embarked
```{r}
prediction <- predict(glmModel, testModImp, type = "raw")

solution <- data.frame(PassengerID = test$PassengerId, Survived = prediction)

solution
write.csv(solution, file = 'glm_solution_2.csv', row.names = F)
```

